---
title: "Milestone 1"
output: pdf_document
---
**Task 1: Choose your favourite dataset**
 1) vancouver_trees
 2) flow_sample
 3) cancer_sample
 4) building_permits
 
*Narrowing down my selection:*
1) vancouver_trees
```{r}
dplyr::summarise(vancouver_trees)
dplyr::select(vancouver_trees, everything())
class(vancouver_trees)
```
2) flow_sample
```{r}
summarise(flow_sample)
select(flow_sample, everything())
class(flow_sample)
```
3) cancer_sample
```{r}
summarise(cancer_sample)
select(cancer_sample, everything())
class(cancer_sample)
```
4)building_permits
summarise(building_permits)
select(building_permits, everything())
class(building_permits)

*My selection:*
My top two choices are 'flow_sample' and 'vancouver_trees'. In hindsight, 'flow_sample' will be an excellent data set to work with as there is some relevance to my own research. My second choice will be'vancouver_trees' due to my longstanding interest in plant biology. Therefore, I will be working with 'flow_sample' due to the relevance with my own research. I will aim to examine changes in flow through time. 


